---
title: "p8105_hw1_jt3649"
author: "Juan Tang"
date: "2025-09-20"
output: html_document
---
## Problem 1
```{r}
library(moderndive)
library(tidyverse)
data("early_january_weather")
```

### Dataset Description
The `early_january_weather` dataset has `r nrow(early_january_weather)` rows (observations) and 
`r ncol(early_january_weather)` columns (variables).
Variables include: origin, date/time variables (year, month, day, hour), temperature in Fahrenheit (temp), dew point temperature (dewp), humidity (humid), wind characteristics variables (wind_air, wind_speed, wind_gust), precipitation (precip), pressure (pressure), visibility (visib), and time_hour. 

The mean temperature across all observations is `r mean(pull(early_january_weather, "temp"), na.rm = TRUE)` ºF.

### Scatterplot of temp vs time_hour
Code for Scatterplot
```{r}
ggplot(early_january_weather,
       aes(x = time_hour, 
           y = temp, 
           color = humid
       )) + geom_point() + labs(title = "Temperature vs. Time Scatterplot", 
             x = "Time(Hour)", 
             y = "Temperature (ºF)", 
             color = "Humidity (%)")
```
Saving the plot
```{r}
ggsave("temperature_vs_time_hour_Scatterplot.png", width = 8, height = 6)
```

### Description of Scatterplot
The scatterplot shows how temperature in New York changed during the first two weeks of January 2013, with x-axis representing time, y-axis representing temperature in ºF, and point color representing humidity. 
The pattern shows a regular daily fluctuation in temperature, presenting a consistent wave-like pattern.The temperature rises during daylight hours and falls at night. Across the two-week period, there appears to be a general upward trend in temperatures, with later days showing higher temperatures overall compared to the beginning of the period. Around January 14, there is a noticeable peak where the temperature exceeded 56°F.

There is an inverse relationship between temperature and humidity in general although the relationship is not strictly linear. Higher temperatures are shown to correspond with lower humidity levels (darker color) and lower temperatures correspond with higher humidity levels (lighter colored points). 


## Problem 2
Creating a dataframe
```{r}
df=tibble(
  random_sample = rnorm(10), 
  vec_positive = random_sample > 0, 
  char_vec = letters[1:10], 
  factor_vec = factor(rep(c("low", "medium", "high"), 
                          length.out = 10),
                      levels = c("low", "medium", "high"))
)
```
This is the dataframe for problem 2: 
```{r}
df
```
Take the mean of each variable in the dataframe df
```{r}
mean_random_sample = mean(pull(df, "random_sample")) 
cat("Mean of random sample variable =", mean_random_sample)

mean_vec_positive = mean(pull(df, "vec_positive"))
cat("Mean of positive vectors =", mean_vec_positive)

mean_char_vec = mean(pull(df, "char_vec"))

mean_factor_vec = mean(pull(df, "factor_vec"))
```

### Explanation
The mean of random_sample is `r mean(pull(df, "random_sample"))`. This mean works because it's numeric value. 
The mean of vec_positive is `r mean(pull(df, "vec_positive"))`. This mean also works because logical values are coerced to 0 (False) or 1 (TRUE). When a mean is requested, the system automatically convert each TRUE to a 1 and each FALSE to a 0. It then adds up all the 1s and 0s, and devides the sum by the total number of values in the list. 
char_vec and factor_vec cannot take mean values because they are text and categorical labels. 

### as.numeric Function
```{r}
as.numeric(pull(df, "random_sample"))
as.numeric(pull(df, "vec_positive"))
as.numeric(pull(df, "char_vec"))
as.numeric(pull(df, "factor_vec"))
```
