---
title: "p8105_hw1_jt3649"
author: "Juan Tang"
date: "`r Sys.Date()`"
output: html_document
---
## Problem 1
```{r}
library(moderndive)
library(tidyverse)
data("early_january_weather")
```

### Dataset Description
The `early_january_weather` dataset has `r nrow(early_january_weather)` rows (observations) and 
`r ncol(early_january_weather)` columns (variables).
Variables include: origin, date/time variables (year, month, day, hour), temperature in Fahrenheit (temp), dew point temperature (dewp), humidity (humid), wind characteristics variables (wind_air, wind_speed, wind_gust), precipitation (precip), pressure (pressure), visibility (visib), and time_hour. 

The mean temperature across all observations is `r mean(pull(early_january_weather, "temp"), na.rm = TRUE)` ºF.

### Scatterplot of temp vs time_hour
Code for Scatterplot
```{r}
ggplot(early_january_weather,
       aes(x = time_hour, 
           y = temp, 
           color = humid
       )) + geom_point() + labs(title = "Temperature vs. Time Scatterplot", 
             x = "Time(Hour)", 
             y = "Temperature (ºF)", 
             color = "Humidity (%)")
```
Saving the plot
```{r}
ggsave("temperature_vs_time_hour_Scatterplot.png", width = 8, height = 6)
```

### Description of Scatterplot
The scatterplot shows how temperature in New York changed during the first two weeks of January 2013, with point color representing humidity. Overall, the temperature gradually increased across the period, but there were daily fluctuations that reflect natural day–night cycles. Around January 14, there is a noticeable peak where the temperature exceeded 55°F. Humidity varies throughout the time series, with lighter-colored points indicating higher humidity. In some cases, higher temperatures coincide with higher humidity, but the relationship is not strictly linear.

Inverse relationship
- higher temperatures (peaks of the waves) correspond with lower humidity levels
- lower temperatures (troughs of the wav) correspond with higher humidity levels (darker colored points)



## Problem 2
Creating a dataframe
```{r}
df=tibble(
  random_sample = rnorm(10), 
  vec_positive = random_sample > 0, 
  char_vec = letters[1:10], 
  factor_vec = factor(rep(c("low", "medium", "high"), 
                          length.out = 10),
                      levels = c("low", "medium", "high"))
)
```
This is the dataframe for problem 2: 
```{r}
df
```
Take the mean of each variable in the dataframe df
```{r}
mean_random_sample = mean(pull(df, "random_sample")) 
cat("Mean of random sample variable =", mean_random_sample)

mean_vec_positive = mean(pull(df, "vec_positive"))
cat("Mean of positive vectors =", mean_vec_positive)

mean_char_vec = mean(pull(df, "char_vec"))

mean_factor_vec = mean(pull(df, "factor_vec"))
```

### Explanation
The mean of random_sample is `r mean(pull(df, "random_sample"))`. This mean works because it's numeric value. 
The mean of vec_positive is `r mean(pull(df, "vec_positive"))`. This mean also works because logical values are coerced to 0 (False) or 1 (TRUE). When a mean is requested, the system automatically convert each TRUE to a 1 and each FALSE to a 0. It then adds up all the 1s and 0s, and devides the sum by the total number of values in the list. 
char_vec and factor_vec cannot take mean values because they are text and categorical labels. 

### as.numeric Function
```{r}
as.numeric(pull(df, "random_sample"))
as.numeric(pull(df, "vec_positive"))
as.numeric(pull(df, "char_vec"))
as.numeric(pull(df, "factor_vec"))
```
